\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-11-28 chapters/bttda.tex
Layout chapter wcble

\chapter{Block-term tensor discriminant analysis}
\label{sec:bttda}
\input{chapters/bttda/tensorviz.tex}
\epigraph{%
  ``I journeyed to the market to purchase a cow, its price set at ninety-and-seven coins.
  Lacking this sum, I therefore borrowed fifty coins from each of two good friends.
  With these hundred coins, I paid for the beast and received three coins in change.
  Of the three, I returned one coin each to my companions, keeping one for myself.
  Now, I owe each friend forty-and-nine coins, a total debt of
  ninety-and-eight. And with mine one remaining, the
  total merely equals ninety-and-nine coins, short of the hundred I borrowed!
  Tell me then, kind stranger, whither hath the missing coin gone?''
}{%
  Ancient Algerian riddle
}
\publicationnote{This chapter was submitted as~\textcite{VanDenKerchove2024a}.}

@@

Additionally, the forward modeling step inherent to BTTDA results
in an interpretable model since we can use the activation patterns or the
forward projection to inspect the neural patterns corresponding to the
relevant discriminatory information at each block~\cite{Haufe2014}.
\Cref{fig:forward} shows the activation patterns
of two blocks obtained from the BNCI2014-008 dataset as well as the forward
projection of the difference between the averages of the mean latent tensor per
class (\emph{contrasts}) after forward projection.
While the weights of the backward projection are
uninterpretable~\cite{Haufe2014},
the activation patterns and contrasts after forward projection clearly show
that ERP components can be recognized and separated into different
BTTDA blocks.
\begin{figure*}[t]
  \begin{subfigure}[b]{\linewidth}
    \begin{minipage}[b]{.15\linewidth}
      \inputpgf{figures/bttda}{forward_block-0_ap-sp-0.pgf}
      \vspace{3em}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.15\linewidth}
      \inputpgf{figures/bttda}{forward_block-0_ap-sp-1.pgf}
      \vspace{3em}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.25\linewidth}
      \inputpgf{figures/bttda}{forward_block-0_ap-tmp.pgf}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.3\linewidth}
      \includegraphics[width=\linewidth]{figures/bttda/forward_block-0_contrast.png}
    \end{minipage}
    \vspace{-1em}
    \subcaption{The first block.}
  \end{subfigure}

  \begin{subfigure}[b]{\linewidth}
    \begin{minipage}[b]{.15\linewidth}
      \inputpgf{figures/bttda}{forward_block-1_ap-sp-0.pgf}
      \vspace{3em}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.15\linewidth}
      \inputpgf{figures/bttda}{forward_block-1_ap-sp-1.pgf}
      \vspace{3em}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.25\linewidth}
      \inputpgf{figures/bttda}{forward_block-1_ap-tmp.pgf}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.3\linewidth}
      \includegraphics[width=\linewidth]{figures/bttda/forward_block-1_contrast.png}
    \end{minipage}
    \vspace{-1em}
    \subcaption{The second block.}
  \end{subfigure}

>>>>

\chapter{Block-term tensor discriminant analysis}
\label{sec:bttda}
\input{chapters/bttda/tensorviz.tex}
\epigraph{
``Traveling through hyperspace ain't like dusting crops, boy!
Without precise calculations we could fly right through a star or bounce too
close to a supernova and that'd end your trip real quick, wouldn't it?''
}{Han Solo in \emph{Star Wars Episode IV: A New Hope}}
\publicationnote{This chapter was submitted as~\textcite{VanDenKerchove2024a}.}

@@

Additionally, the forward modeling step inherent to BTTDA results
in an interpretable model since we can use the activation patterns or the
forward projection to inspect the neural patterns corresponding to the
relevant discriminatory information at each block~\cite{Haufe2014}.
\Cref{fig:forward} shows the activation patterns
of two blocks obtained from the BNCI2014-008 dataset as well as the forward
projection of the difference between the averages of the mean latent tensor per
class (\emph{contrasts}) after forward projection.
While the weights of the backward projection are
uninterpretable~\cite{Haufe2014},
the activation patterns and contrasts after forward projection clearly show
that ERP components can be recognized and separated into different
BTTDA blocks.
\begin{figure*}[t]
  \begin{subfigure}[b]{\linewidth}
    \begin{minipage}[b]{.15\linewidth}
      \inputpgf{figures/bttda}{forward_block-0_ap-sp-0.pgf}
      \vspace{3em}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.15\linewidth}
      \inputpgf{figures/bttda}{forward_block-0_ap-sp-1.pgf}
      \vspace{3em}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.25\linewidth}
      \inputpgf{figures/bttda}{forward_block-0_ap-tmp.pgf}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.3\linewidth}
      \includegraphics[width=\linewidth]{figures/bttda/forward_block-0_contrast.png}
    \end{minipage}
    %\vspace{-1em}
    \subcaption{The first block.}
  \end{subfigure}

  \begin{subfigure}[b]{\linewidth}
    \begin{minipage}[b]{.15\linewidth}
      \inputpgf{figures/bttda}{forward_block-1_ap-sp-0.pgf}
      \vspace{3em}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.15\linewidth}
      \inputpgf{figures/bttda}{forward_block-1_ap-sp-1.pgf}
      \vspace{3em}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.25\linewidth}
      \inputpgf{figures/bttda}{forward_block-1_ap-tmp.pgf}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.3\linewidth}
      \includegraphics[width=\linewidth]{figures/bttda/forward_block-1_contrast.png}
    \end{minipage}
    %\vspace{-1em}
    \subcaption{The second block.}
  \end{subfigure}

### 2024-11-28 chapters/stbf_struct.tex
Layout chapter stbf-struct
Bleed
Layout BTTDA
Layout chapter wcble

\chapter{Kronecker-structured linear decoding}
\label{sec:stbf-struct}
\newcommand{\pv}[1]{
	%\ifdim#1pt<0.05pt
	%	\cellcolor{black!20}
	%\fi
	%\ifdim#1pt<0.001pt
	%	$< 0.001$
	%\else
	%	\ifdim#1pt=1.000pt
	%		\multicolumn{1}{c}{--}
	%	\else
	%		$#1$
	%	\fi
	%\fi
	#1
}
% Kronecker-structured LDA ====================================================

@@

  \begin{table}[h!]
      \small
      \sffamily
	    \input{chapters/stbf_struct-acc-stbf_kron_toep_shrunk.tex}
      \caption[Accuracies for \acs{stbf-struct}.]{Accuracies (\%) for \ac{stbf-struct}.}
      \label{tab:stbf-struct/accuracies/stbf-struct}
  \end{table}
  \begin{table}[h!]
      \small
      \sffamily
	    \input{chapters/stbf_struct-acc-stbf_loocv.tex}
      \caption[Accuracies for \acs{stbf-shrunk}.]{Accuracies (\%) for
      \ac{stbf-shrunk}.}
      \label{tab:stbf-struct/accuracies/stbf-shrunk}
  \end{table}
  \begin{table}[h!]
      \small
      \sffamily
	    \input{chapters/stbf_struct-acc-stbf_emp.tex}
      \caption[Accuracies for \acs{stbf-emp}.]{Accuracies (\%) for
      \ac{stbf-emp}.}
      \label{tab:stbf-struct/accuracies/stbf-emp}
  \end{table}
  \begin{table}[h!]
      \small
      \sffamily
	    \input{chapters/stbf_struct-acc-xdawn_rg.tex}
      \caption[Accuracies for XDAWN+RG.]{Accuracies (\%) for
      XDAWN+RG.}
      \label{tab:stbf-struct/accuracies/xdawn-rg}
  \end{table}

@@

  \fullpagefig{%
    \input{figures/stbf_struct/accuracy_all.pgf}
  }{%
    \caption[Accuracy averaging over different numbers of trials.]{%
      Accuracy of the different classifiers for all 21 subjects relative to the
		  number of blocks available for training. One block consists of 135
		  epochs and corresponds to 27 seconds of stimulation. Accuracies
		  are shown for the evaluation settings averaging over different
	    numbers of trials, ranging from 1 to 15.
    }
    \label{fig:stbf-struct/accuracy-ap}
  }

  \clearpage
  \section{Statistical comparisons}
  \begin{table}[h]
    \footnotesize
    \begin{subtable}{\linewidth}
	    \input{chapters/stbf_struct_tests_1.tex}
	    \caption{Evaluation on one testing trial.}
      \label{tab:stbf-struct/p-values-1}
    \end{subtable}
    \smallskip

    \begin{subtable}{\linewidth}
	    \input{chapters/stbf_struct_tests_2.tex}
	    \caption{Evaluation averaging over two testing trials.}
	    \label{tab:stbf-struct/p-values-2}
    \end{subtable}
    \smallskip

    \begin{subtable}{\linewidth}
	    \input{chapters/stbf_struct_tests_3.tex}
	    \caption{Evaluation averaging over five testing trials.}
	    \label{tab:stbf-struct/p-values-5}
    \end{subtable}
    \caption[Statistical significance of differences in classifier
    performance.]{%
        Statistical significance of differences in classifier performance.
        $p$-values calculated by one-sided Wilcoxon signed-rank test with
	    	Holm correction using one testing trial for different classifiers
	    	and levels of data availability. $p$-values $<0.05$ are considered
        significant.
    }
	  \label{tab:stbf-struct/p-values}
	\end{table}

>>>>

\chapter{Kronecker-structured linear decoding}
\label{sec:stbf-struct}
\newcommand{\pv}[1]{
	%\ifdim#1pt<0.05pt
	%	\cellcolor{black!20}
	%\fi
	%\ifdim#1pt<0.001pt
	%	$< 0.001$
	%\else
	%	\ifdim#1pt=1.000pt
	%		\multicolumn{1}{c}{--}
	%	\else
	%		$#1$
	%	\fi
	%\fi
	#1
}
\epigraph{%
  ``I journeyed to the market to purchase a cow, its price set at ninety-and-seven coins.
  Lacking this sum, I therefore borrowed fifty coins from each of two good friends.
  With these hundred coins, I paid for the beast and received three coins in change.
  Of the three, I returned one coin each to my companions, keeping one for myself.
  Now, I owe each friend forty-and-nine coins, a total debt of
  ninety-and-eight. And with mine one remaining, the
  total merely equals ninety-and-nine coins, short of the hundred I borrowed!
  Tell me then, kind stranger, whither hath the missing coin gone?''
}{%
  Ancient Algerian riddle
}
% Kronecker-structured LDA ====================================================

@@

  \begin{table}[!ht]
      \small
      \sffamily
	    \input{chapters/stbf_struct-acc-stbf_kron_toep_shrunk.tex}
      \caption[Accuracies for \acs{stbf-struct}.]{Accuracies (\%) for \ac{stbf-struct}.}
      \label{tab:stbf-struct/accuracies/stbf-struct}
  \end{table}
  \begin{table}[!ht]
      \small
      \sffamily
	    \input{chapters/stbf_struct-acc-stbf_loocv.tex}
      \caption[Accuracies for \acs{stbf-shrunk}.]{Accuracies (\%) for
      \ac{stbf-shrunk}.}
      \label{tab:stbf-struct/accuracies/stbf-shrunk}
  \end{table}
  \begin{table}[!ht]
      \small
      \sffamily
	    \input{chapters/stbf_struct-acc-stbf_emp.tex}
      \caption[Accuracies for \acs{stbf-emp}.]{Accuracies (\%) for
      \ac{stbf-emp}.}
      \label{tab:stbf-struct/accuracies/stbf-emp}
  \end{table}
  \begin{table}[!ht]
      \small
      \sffamily
	    \input{chapters/stbf_struct-acc-xdawn_rg.tex}
      \caption[Accuracies for XDAWN+RG.]{Accuracies (\%) for
      XDAWN+RG.}
      \label{tab:stbf-struct/accuracies/xdawn-rg}
  \end{table}

@@

  \fullpagefig[caption.south, anchor.north]{%
    \input{figures/stbf_struct/accuracy_all.pgf}
  }{%
    \caption[Accuracy averaging over different numbers of trials.]{%
      Accuracy of the different classifiers for all 21 subjects relative to the
		  number of blocks available for training. One block consists of 135
		  epochs and corresponds to 27 seconds of stimulation. Accuracies
		  are shown for the evaluation settings averaging over different
	    numbers of trials, ranging from 1 to 15.
    }
    \label{fig:stbf-struct/accuracy-ap}
  }

  \clearpage
  \section{Statistical comparisons}
  \vfill
  \begin{table}[!ht]
    \centering
    \begin{adjustbox}{angle=90}
      \begin{subtable}{\linewidth}
	      \input{chapters/stbf_struct_tests_1.tex}
	      \caption{Evaluation on one testing trial.}
        \label{tab:stbf-struct/p-values-1}
      \end{subtable}
    \end{adjustbox}
    \begin{adjustbox}{angle=90}
      \begin{subtable}{\linewidth}
	      \input{chapters/stbf_struct_tests_2.tex}
	      \caption{Evaluation averaging over two testing trials.}
	      \label{tab:stbf-struct/p-values-2}
      \end{subtable}
    \end{adjustbox}
    \begin{adjustbox}{angle=90}
      \begin{subtable}{\linewidth}
	      \input{chapters/stbf_struct_tests_3.tex}
	      \caption{Evaluation averaging over five testing trials.}
	      \label{tab:stbf-struct/p-values-5}
      \end{subtable}
    \end{adjustbox}
    \caption[Statistical significance of differences in classifier
    performance.]{%
        Statistical significance of differences in classifier performance.
        $p$-values calculated by one-sided Wilcoxon signed-rank test with
	    	Holm correction using one testing trial for different classifiers
	    	and levels of data availability. $p$-values $<0.05$ are considered
        significant.
    }
	  \label{tab:stbf-struct/p-values}
	\end{table}
  \vfill
